---
title: "Exam"
format: html
editor: visual
---

## Exercise 1

### a

Each of the individual common bean seed is the sampling unit.

### b

The germination times is continuous.

### c

- Dependent variables is germination times

- Independent variable is the moisture level

### d

The sample size is 90 seeds

### e

1. We can't use ANOVA test because the data is bimodal and not normally distributed.

2. Yes we can, due to the transformation which makes the data normally distributed.

3. Kruskal wallis test can be used since it is the non-parametric counterpart of the ANOVA test.

4. You can't use t-test because the data has three group levels.

5. Yes you can visualize the data using boxplots to better understand group differences after choosing a statistical test. 

## Exercise 2

## a

```{r}

df1 <- read.csv("envdata-20250107data1.csv")
View(df1)
```

## b
```{r}
head(df1)

str(df1)
```

### c

WaterPH because water with high acidity or alkalinity can be harmful for consumption making it a key measurement to measure water PH level.

### d


